Analysis of finite-volume advection schemes on cubed-sphere grids and an accurate \textcolor{red}{scheme} for divergent winds

### 2024-05-20 paper.tex
several minor corrections

The transport scheme employed in FV3 extends the method introduced by \cite{lin:1996} to the cubed-sphere, as proposed by \cite{putman:2007}. 
\textcolor{red}{
	The major challenge of transitioning from latitude-longitude grids to grids based on Platonic solids \cite{stan:2012} is grid imprinting, which refers to the appearance of grid features on the solution (cf. e.g., \cite{peixoto:2013, weller:2012, mouallem:2023}), and is highly undesirable due to its lack of physical meaning.}

@@

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cubed-sphere grids}
\label{cs-grids}
The goal of this section is to review the concept of cubed-sphere grids, particularly those available in FV3.
To start with, we are going to present the mapping between the cube and the sphere introduced by \cite{sadourny:1972} in section \ref{cs-equidistant}, also known as the equidistant mapping.
Following that, we show that using a change of coordinates, we may use the equidistant mapping to create other mappings from the cube to the sphere.
Namely, section \ref{cs-equiangular} introduces the equiangular mapping introduced by \cite{ronchi:1996}, and section \ref{cs-equiedge} introduces the equi-edge mapping introduced by \cite{chen:2021}. 
In section \ref{cs-grid}, we demonstrate how these mappings are utilized to generate cubed-sphere grids. 
We mainly introduce all the notations and tools that will be needed for the remainder of this work.
Finally, in section \ref{cs-ghost}, we demonstrate how the ghost cells that we will need are generated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Equidistant mapping}
\label{cs-equidistant}
The mapping introduced by \cite{sadourny:1972} between the cube and the sphere divides the sphere into six quadrilaterals, also called panels, and enables the tessellation of the sphere into smaller quadrilaterals for each panel.
This mapping is also called equidistant mapping and generates the equidistant grid.
Given $R>0$, the sphere of radius $R$ 
centered at the origin of  $\mathbb{R}^3$ is denoted as:
\begin{equation}
	\label{s2_r}
	\mathbb{S}^2_R = \{ P = (p_x,p_y,p_z) \in \mathbb{R}^3: p_x^2 + p_y^2 + p_z^2 = R^2\}.
\end{equation}
For the purposes of this work, the Earth radius $R=6.371\times 10^6$ meters is considered.
The equidistant mapping considers a cube centered at the origin with a side length of $\frac{2R}{\sqrt{3}}$ and radially projects the cube faces onto the sphere (Figure \ref{sph-cube-equidist}).

The equidistant mapping is a family of maps ${\Gamma}_{p}: [-1,1] \times [-1,1] \to \mathbb{S}^2_R$, where $p=1, \ldots, 6$, defined as follows:
\begin{align}
	\label{gamma_1}
	{\Gamma}_{1}(X,Y) &= \frac{R}{\sqrt{1 + X^2 + Y^2}}(1, X, Y),\\ 
	\label{gamma_2}
	{\Gamma}_{2}(X,Y) &= \frac{R}{\sqrt{1 + X^2 + Y^2}}(-X, 1, Y), \\
	\label{gamma_3}
	{\Gamma}_{3}(X,Y) &= \frac{R}{\sqrt{1 + X^2 + Y^2}}(-X, -Y, 1), \\
	\label{gamma_456}
{\Gamma}_{3+k}(X,Y)&=-{\Gamma}_{k}(Y,X),\quad k=1,2,3.
\end{align}
These mappings allow for the coverage of the sphere and represent the relationship between the sphere and the cube.
The idea behind this mapping is illustrated in Figure \ref{sph-cube-equidist}, where the coordinates $(X,Y)$ are thought to live on the cube faces.
Figure \ref{sph-cube-equidist} shows how the grid points on the sphere are equally spaced on the cube and then projected onto the sphere, hence the name equidistant.
Note that there are other ways to arrange the coordinates over the panels.
Each one defines an interlock pattern, as discussed by \cite[Section 2.1]{chen:2021}.
The arrangement presented here is known as the staircase arrangement, which is used in FV3 and provides some advantages for exchanging information between panels.
\begin{figure}[!htb]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=1.1\linewidth]{g1}
		\caption{Cube and sphere equidistant mapping for $p_z=0$.\label{sph-cube-equidist}}
	\end{subfigure}

@@

The derivative of ${\Gamma}_p$ is a $3\times2$ matrix denoted by $d{\Gamma}_p$. Explicit formulas are provided in  \ref{metric_terms}.
With the aid of the derivative,  a basis of tangent vectors $\{{\partial_X {\Gamma}_p, \partial_Y {\Gamma}_p}\}$ may be defined
at each point on the sphere, where $\partial_X {\Gamma}_p$ is given by the first column of $d{\Gamma}_p$ 
and $\partial_Y {\Gamma}_p$ is given by the second column of $d{\Gamma}_p$.
Along with, the metric tensor is defined as
$G_{{\Gamma}} := (d\Gamma_p)^T \cdot d\Gamma_p$.
It is easy to see that the metric tensor does not depend on $p$.
The Jacobian of the metric tensor $G_{{\Gamma}}$ is then defined as $\sqrt{\mathfrak{g}_{{\Gamma}}} :=
\sqrt{|\det{G_{{\Gamma}}}|}.$

@@

\textcolor{red}{Given a tangent vector field $\boldsymbol{u}$ on the sphere, also known as wind, we may represent it using the basis obtained by cubed-sphere coordinates:
\begin{equation}
	\label{contravariant-wind}
	\boldsymbol{u}(x, y, p) = 
	\mathfrak{u}(x, y,p ) \partial_x{\Psi}_p(x, y) + 
	\mathfrak{v}(x, y, p) \partial_y{\Psi}_p(x, y).
\end{equation}
This representation $(\mathfrak{u},\mathfrak{v})$ is known as the contravariant representation, and they are the components of the wind on the tangent basis defined by the cubed-sphere mapping.
A detailed discussion on how the cubed-sphere wind representation is related to the zonal and meridional representation is presented in \ref{ap-wind}.
In practice, FV3 schemes \cite{putman:2007, harris:2021} use the normalized contravariant wind
$({u},{v})$ given by:
\begin{equation}
	\label{norm-contravariant-wind}
	\boldsymbol{u}(x, y, p) = 
	{u}(x, y, p) \boldsymbol{e}_x(x, y, p) + 
	{v}(x, y, p) \boldsymbol{e}_y(x, y, p),
\end{equation}
where $\boldsymbol{e}_x$ and $\boldsymbol{e}_y$ are the normalized cubed-sphere tangent vectors, which may be computed easily in terms of the grid points \cite[Appendix C2]{chen:2021}.
It is easy to see that:
\begin{equation}
	\label{contra-uv}
	\mathfrak{u}(x,y,p)  = \frac{{u}(x,y,p)}{\|\partial_x{\Psi}_p(x,y)\|}, \quad
	\mathfrak{v}(x,y,p)  = \frac{{v}(x,y,p)}{\|\partial_y{\Psi}_p(x,y)\|}.
\end{equation}
Finally, we recall that the horizontal divergence operator for a  wind $\boldsymbol{u}$  on the sphere is defined in terms of the cubed-sphere metric terms as follows:
\begin{align}
	\label{div-def}
	[\nabla \cdot {\boldsymbol{u}}](x,y,p) :=
	\frac{1}{\sqrt{\mathfrak{g}_{\Psi}}(x,y)}
	\bigg(
	\partial_x(\sqrt{\mathfrak{g}_{\Psi}}\mathfrak{u})(x,y,p) +
	\partial_y(\sqrt{\mathfrak{g}_{\Psi}}\mathfrak{v})(x,y,p)
	\bigg),
\end{align}
for $x,y\in[-a,a]$, $p$ is the panel and
$\mathfrak{u}$ and $\mathfrak{v}$ are the contravariant wind components.
The divergence operator shall be used in the transport model in Section \ref{adv-2d}.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Equiangular mapping}
\label{cs-equiangular}
Another cubed-sphere mapping is the equiangular mapping, 
introduced by \cite{ronchi:1996}, which leads to a more uniform grid.
This mapping is obtained by considering $\beta(x) = \tan{x}$ and $a=\frac{\pi}{4}$.
In this case, $\beta(x)$ represents the angular coordinates, and the cube-sphere is obtained by partitioning the angle between
grid points equally, as illustrated in Figure \ref{sph-cube-equiangular}, hence the name equiangular.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cubed-sphere grid generation}
\label{cs-grid}
Let us fix two positive integers $N$ and $\nu$, where $N$ represents the number of cells in a direction and $\nu$ represents the number of ghost cell layers.
The equiangular or equi-edge mapping, denoted by $\Psi_p$, introduced previously, is considered to generate the cubed-sphere grid.
For simplicity, the metric term $\sqrt{\mathfrak{g}}_{\Psi}$ is denoted by $\sqrt{\mathfrak{g}}$.
To generate the cubed-sphere grid, the domain $[-a,a]\times[-a,a]$ is discretized using uniformly spaced points.
\begin{align}
\label{edge-coords}
x_{i+\frac{1}{2}} = -a+i\Delta x, \quad
y_{j+\frac{1}{2}} = -a+j\Delta y,
\end{align}
where $\Delta x = \Delta y = \frac{2a}{N}$, $i,j=-\nu+1,\ldots,N+1+\nu$. The center coordinates  are defined as:
\begin{align}
	\label{center-coords}
	x_{i} = \frac{x_{i+\frac{1}{2}}+x_{i-\frac{1}{2}}}{2}, \quad
	y_{j} = \frac{y_{j+\frac{1}{2}}+y_{j-\frac{1}{2}}}{2},
\end{align}
for $i,j=-\nu+1,\ldots,N+\nu$.
Notice that the mappings ${\Psi}_p$ defined before can be computed outside the range $[-a,a]$, and the cubed-sphere mapping can be applied to all these ghost cell points.
Firstly, we shall focus the attention on the interior cells; the generation of ghost cells shall be addressed in Section \ref{cs-ghost}.

@@

There are four types of grid points on the cubed-sphere that are needed to be computed: the center, corners, right-left edge midpoints, and  up-down edge midpoints. 
\textcolor{red}
{These points are illustrated in Figure \ref{cgrid}.}
The corner points are computed as:
\begin{equation}
	\label{corner-points}
	\Psi_{i+\frac{1}{2},j+\frac{1}{2},p} := {\Psi}_p(x_{i+\frac{1}{2}},y_{j+\frac{1}{2}}),
\end{equation}
$i,j=0, \ldots, N$.
To ease the notation hereafter, the dependence on $p$ is omitted because it does not interfere with what is going to be discussed in this section.
Figure \ref{cs-grids-N10} shows the obtained grid lines in for $N=10$.
\begin{figure}[!htb]
	\begin{subfigure}{0.42\textwidth}
		\centering
		\includegraphics[width=1.1\linewidth]{gnomonic_equiedge_cs_10_sphere}
		\caption{Equiangular grid}
	\end{subfigure}
	\centering
\begin{subfigure}{0.42\textwidth}
	\centering
	\includegraphics[width=1.1\linewidth]{gnomonic_equiangular_cs_10_sphere}
	\caption{Equi-edge grid}
\end{subfigure}
	\caption{(a) Illustration of the resulting gridlines
		for the cubed-sphere equiangular and equi-edge mapping for $N=10$.\label{cs-grids-N10}}
\end{figure}

@@

\section{The conservative transport equation on the cubed-sphere}
\label{adv-2d}
The goal of this section is to present and solve numerically the conservative transport equation on the cubed-sphere. 
We are going to consider the equi-edge or equiangular cubed-sphere mappings $\Psi_p$ and their respective local coordinates $(x,y)$.
\textcolor{red}{Once again, the metric term is denoted by $\sqrt{\mathfrak{g}}_{\Psi}$ by $\sqrt{\mathfrak{g}}$ for simplicity.}
Following \cite{nair:2010}, the transport model without sources or sinks is considered:
\begin{align}
	\label{transp-rho}
	[\partial_t \rho + \nabla \cdot (\rho \boldsymbol{u})](x,y,p,t)&=0,\\
	\label{transp-phi}
	[\partial_t (\rho \phi) + \nabla \cdot (\rho \phi \boldsymbol{u})](x,y,p,t)&=0,
\end{align}
for $x,y\in[-a,a]$, $t \in [0,T]$, where $T$ is the final time,
\textcolor{red}
{$\boldsymbol{u}$ is the wind, $\mathfrak{u}$ and $\mathfrak{v}$ are the contravariant wind components (Equation \ref{contravariant-wind}),
$\rho$ is the fluid density and $\phi$ is the tracer concentration}.
Additionally, for the transport model, it is assumed that the initial conditions are given as $\rho(x,y,p,0)=\rho_0(x,y,p)$ and $\phi(x,y,p,0)=\phi_0(x,y,p)$, given $\rho_0$ and $\phi_0$. % and that the boundary conditions are related to the cubed-sphere topology.
Equation \eqref{transp-rho} is the continuity equation and Equation \eqref{transp-phi} is the conservative advection equation.
In this framework, $\rho \phi$ represents the tracer density, and the total masses of $\rho$ and $\rho \phi$ are preserved.  Therefore, these quantities are referred to as conserved quantities.

@@

To concluded this section, the transport model described by Equations \eqref{transp-rho} and \eqref{transp-phi} may be solved using the scheme from Equation \eqref{tr-scheme}, applied considering $q^n=\rho^n$ and $q^n=(\rho \phi)^n$ simultaneously.
In this framework, the tracer concentration is given by $\phi^n = \frac{(\rho \phi)^n}{\rho^n}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical experiments}
\label{num-exp}
\textcolor{red}{
In this section, the goal is to present simulations of the numerical solution of the transport model, governed by Equations \eqref{transp-rho} and \eqref{transp-phi}, utilizing the FV3 and LT2 schemes. As mentioned earlier, the tracer concentration $\phi$ is advected in the transport model.}
To specify the simulation, including the initial condition for the tracer concentration $\phi_0$, the zonal wind component  denoted by $u_\lambda$, and the meridional wind component denoted by $v_\theta$, need to be defined.
The conversion from these wind components to cubed-sphere contravariant wind components (Equation \eqref{contravariant-wind}) is detailed in \ref{ap-wind}.

@@

\subsection{Rotated zonal wind experiments}
In this section, the following rotated zonal wind field based on \cite{will:1992} is considered:
\begin{align}
	\label{zonal-wind}
	\begin{cases}
		u_\lambda(\lambda,\theta,t) = u_0(\cos(\theta)\cos(\alpha) + \sin(\theta)\cos(\lambda)\sin(\alpha)),\\
		v_\theta(\lambda,\theta,t) = -u_0\sin(\lambda)\sin(\alpha),
	\end{cases}
\end{align}
where $\lambda \in [-\pi,\pi]$ represents longitude and $\theta \in [-\frac{\pi}{2},\frac{\pi}{2}]$ represents latitude. 
It is easy to see that the wind is divergence free.
The parameter $\alpha$ is the rotation angle. Following \cite{putman:2007}, the rotation angle is set as $\alpha=\frac{\pi}{4}$ so that the wind is oriented with the cube corners. 
Finally, the parameter $u_0$ is defined as $u_0 = \frac{2\pi R}{T}$.
With this choice, the simulation period is 12 days. The solution should converge to the initial condition after this period, enabling us to compute the final error.
Therefore, the temporal evolution of the error may be analyzed.
For an expression of the exact solution in this case and a general initial condition, refer to \cite[Theorem 5.1, p. 155]{brachet:2018}.
The time-step for $N=48$ is given by $\Delta t^{(0)} = 3600$ seconds, leading to a CFL number of approximately 0.95.

>>>>

The transport scheme employed in FV3 extends the method introduced by \cite{lin:1996} to the cubed-sphere, as proposed by \cite{putman:2007}. 
\textcolor{red}{
	The major challenge of transitioning from latitude-longitude grids to grids based on Platonic solids \cite{stan:2012} is grid imprinting, which refers to the appearance of grid features on the solution (cf. e.g., \cite{peixoto:2013, weller:2012, mouallem:2023}), and is highly undesirable due to its lack of physical meaning.
GHOSTCELLS}

@@

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cubed-sphere grids}
\label{cs-grids}
The goal of this Section is to briefly review the concept of cubed-sphere grids, particularly those available in FV3.
To start with, we are going to present the mapping between the cube and the sphere introduced by \cite{sadourny:1972} in Section \ref{cs-equidistant}, also known as the equidistant mapping.
Following that, we show that using a change of coordinates, we may use the equidistant mapping to create other mappings from the cube to the sphere.
Namely, Section \ref{cs-equiangular} introduces the equiangular mapping introduced by \cite{ronchi:1996}, and Section \ref{cs-equiedge} introduces the equi-edge mapping introduced by \cite{chen:2021}. 
In Section \ref{cs-grid}, we demonstrate how these mappings are utilized to generate cubed-sphere grids. 
We mainly introduce all the notations and tools that will be needed for the remainder of this work.
Finally, in Section \ref{cs-ghost}, we demonstrate how the ghost cells that we will need are generated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Equidistant mapping}
\label{cs-equidistant}
Considering a cube circumscribed on a sphere, \cite{sadourny:1972} introduces a mapping between the cube faces and the sphere, generating a spherical cube.
This mapping is also called equidistant mapping and generates the equidistant grid.
Given $R>0$, the sphere of radius $R$ 
centered at the origin of  $\mathbb{R}^3$ is denoted as:
\begin{equation}
	\label{s2_r}
	\mathbb{S}^2_R = \{ P = (p_x,p_y,p_z) \in \mathbb{R}^3: p_x^2 + p_y^2 + p_z^2 = R^2\}.
\end{equation}
For the purposes of this work, the Earth radius $R=6.371\times 10^6$ meters is considered.
The equidistant mapping considers a cube centered at the origin with a side length of $\frac{2R}{\sqrt{3}}$ and radially projects the cube faces onto the sphere (Figure \ref{sph-cube-equidist}).

The equidistant mapping is a family of maps ${\Gamma}_{p}: [-1,1] \times [-1,1] \to \mathbb{S}^2_R$, where $p=1, \ldots, 6$, defined as follows:
\begin{align}
	\label{gamma_1}
	{\Gamma}_{1}(X,Y) &= \frac{R}{\sqrt{1 + X^2 + Y^2}}(1, X, Y),\\ 
	\label{gamma_2}
	{\Gamma}_{2}(X,Y) &= \frac{R}{\sqrt{1 + X^2 + Y^2}}(-X, 1, Y), \\
	\label{gamma_3}
	{\Gamma}_{3}(X,Y) &= \frac{R}{\sqrt{1 + X^2 + Y^2}}(-X, -Y, 1), \\
	\label{gamma_456}
{\Gamma}_{3+k}(X,Y)&=-{\Gamma}_{k}(Y,X),\quad k=1,2,3.
\end{align}
\textcolor{red}{
These mappings are defined individually for each of the 6 cube faces, also called panels, denoted by $p$, and they allow for the coverage of the sphere.}
The idea behind this mapping is illustrated in Figure \ref{sph-cube-equidist}, where the coordinates $(X,Y)$ are thought to live on the cube faces.
Figure \ref{sph-cube-equidist} shows how the grid points are equally spaced on the cube and then projected onto the sphere, hence the name equidistant.
Note that there are other ways to arrange the coordinates over the panels.
Each one defines an interlock pattern, as discussed by \cite[Section 2.1]{chen:2021}.
The arrangement presented here is known as the staircase arrangement \textcolor{red}{(see Figure 2 in \cite{chen:2021})}, which is used in FV3 and provides some advantages for exchanging information between panels.
\begin{figure}[!htb]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=1.1\linewidth]{g1}
		\caption{Cube and sphere equidistant mapping for $p_z=0$.\label{sph-cube-equidist}}
	\end{subfigure}

@@

The derivative of ${\Gamma}_p$ is a $3\times2$ matrix denoted by $d{\Gamma}_p$. Explicit formulas are provided in  \ref{metric_terms}.
With the aid of the derivative,  a basis of tangent vectors $\{{\partial_X {\Gamma}_p, \partial_Y {\Gamma}_p}\}$ may be defined
at each point on the sphere, where $\partial_X {\Gamma}_p$ is given by the first column of $d{\Gamma}_p$ 
and $\partial_Y {\Gamma}_p$ is given by the second column of $d{\Gamma}_p$.
Along side, the metric tensor is defined as
$G_{{\Gamma}} := (d\Gamma_p)^T \cdot d\Gamma_p$.
It is easy to see that the metric tensor does not depend on the panel $p$.
The Jacobian of the metric tensor $G_{{\Gamma}}$ is then defined as $\sqrt{\mathfrak{g}_{{\Gamma}}} :=
\sqrt{|\det{G_{{\Gamma}}}|}.$

@@

\textcolor{red}{Given a tangent vector field $\boldsymbol{u}$ on the sphere, also known as wind, we may represent it using the basis obtained by cubed-sphere coordinates:
\begin{equation}
	\label{contravariant-wind}
	\boldsymbol{u}(x, y, p) = 
	\mathfrak{u}(x, y,p ) \partial_x{\Psi}_p(x, y) + 
	\mathfrak{v}(x, y, p) \partial_y{\Psi}_p(x, y).
\end{equation}
This representation $(\mathfrak{u},\mathfrak{v})$ is known as the contravariant representation, and they are the components of the wind on the tangent basis defined by the cubed-sphere mapping.
A detailed discussion on how the cubed-sphere wind representation is related to the zonal and meridional representation is presented in \ref{ap-wind}.
In practice, FV3 schemes \cite{putman:2007, harris:2021} use the normalized contravariant wind
$({u},{v})$ given by:
\begin{equation}
	\label{norm-contravariant-wind}
	\boldsymbol{u}(x, y, p) = 
	{u}(x, y, p) \boldsymbol{e}_x(x, y, p) + 
	{v}(x, y, p) \boldsymbol{e}_y(x, y, p),
\end{equation}
where $\boldsymbol{e}_x$ and $\boldsymbol{e}_y$ are the normalized cubed-sphere tangent vectors, which may be computed easily in terms of the grid points \cite[Appendix C2]{chen:2021}.
It is easy to see that:
\begin{equation}
	\label{contra-uv}
	\mathfrak{u}(x,y,p)  = \frac{{u}(x,y,p)}{\|\partial_x{\Psi}_p(x,y)\|}, \quad
	\mathfrak{v}(x,y,p)  = \frac{{v}(x,y,p)}{\|\partial_y{\Psi}_p(x,y)\|}.
\end{equation}
Finally, we recall that the horizontal divergence operator for a  wind $\boldsymbol{u}$  on the sphere is defined in terms of the cubed-sphere metric terms as follows:
\begin{align}
	\label{div-def}
	[\nabla \cdot {\boldsymbol{u}}](x,y,p) :=
	\frac{1}{\sqrt{\mathfrak{g}_{\Psi}}(x,y)}
	\bigg(
	\partial_x(\sqrt{\mathfrak{g}_{\Psi}}\mathfrak{u})(x,y,p) +
	\partial_y(\sqrt{\mathfrak{g}_{\Psi}}\mathfrak{v})(x,y,p)
	\bigg),
\end{align}
for $x,y\in[-a,a]$, $p$ is the panel and
$\mathfrak{u}$ and $\mathfrak{v}$ are the contravariant wind components.
The divergence operator shall be used in the transport model in Section \ref{adv-2d}.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Equiangular mapping}
\label{cs-equiangular}
Another cubed-sphere mapping is the equiangular mapping, 
introduced by \cite{ronchi:1996}, which leads to a more uniform grid on the sphere.
This mapping is obtained by considering $\beta(x) = \tan{x}$ and $a=\frac{\pi}{4}$.
In this case, $\beta(x)$ represents the angular coordinates, and the cube-sphere is obtained by partitioning the angle between
grid points equally, as illustrated in Figure \ref{sph-cube-equiangular}, hence the name equiangular.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cubed-sphere grid generation}
\label{cs-grid}
Let us fix two positive integers $N$ and $\nu$, where $N$ represents the number of cells in both the $x$ and $y$ directions, and $\nu$ represents the number of ghost cell layers. 
The equiangular or equi-edge mappings, denoted by $\Psi_p$, introduced previously, are considered to generate the cubed-sphere grid.
\textcolor{red}{The notation $\Psi_p$ is used for both equiangular and equi-edge mappings, as what will be discussed does not depend particularly on the mapping.}
For simplicity, the metric term $\sqrt{\mathfrak{g}}_{\Psi}$ is denoted by $\sqrt{\mathfrak{g}}$.
To generate the cubed-sphere grid, the domain $[-a,a]\times[-a,a]$ is discretized using uniformly spaced points.
\begin{align}
\label{edge-coords}
x_{i+\frac{1}{2}} = -a+i\Delta x, \quad
y_{j+\frac{1}{2}} = -a+j\Delta y,
\end{align}
where $\Delta x = \Delta y = \frac{2a}{N}$, $i,j=-\nu+1,\ldots,N+1+\nu$. The center coordinates  are defined as:
\begin{align}
	\label{center-coords}
	x_{i} = \frac{x_{i+\frac{1}{2}}+x_{i-\frac{1}{2}}}{2}, \quad
	y_{j} = \frac{y_{j+\frac{1}{2}}+y_{j-\frac{1}{2}}}{2},
\end{align}
for $i,j=-\nu+1,\ldots,N+\nu$.
Notice that the mappings ${\Psi}_p$ defined before can be computed outside the range $[-a,a]$, and the cubed-sphere mapping can be applied to all these ghost cell points.
Firstly, we shall focus the attention on the interior cells; the generation of ghost cells shall be addressed in Section \ref{cs-ghost}.

@@

There are four types of grid points on the cubed-sphere that are needed to be computed: the center, corners, right-left edge midpoints, and  up-down edge midpoints. 
\textcolor{red}
{These points are illustrated in Figure \ref{cgrid}.}
The corner points are computed as:
\begin{equation}
	\label{corner-points}
	\Psi_{i+\frac{1}{2},j+\frac{1}{2},p} := {\Psi}_p(x_{i+\frac{1}{2}},y_{j+\frac{1}{2}}),
\end{equation}
$i,j=0, \ldots, N$.
To ease the notation hereafter, the dependence on $p$ is omitted because it does not interfere with what is going to be discussed in this Section.
Figure \ref{cs-grids-N10} shows the obtained grid lines in for $N=10$.
\begin{figure}[!htb]
	\begin{subfigure}{0.42\textwidth}
		\centering
		\includegraphics[width=1.1\linewidth]{gnomonic_equiedge_cs_10_sphere}
		\caption{Equiangular grid}
	\end{subfigure}
	\centering
\begin{subfigure}{0.42\textwidth}
	\centering
	\includegraphics[width=1.1\linewidth]{gnomonic_equiangular_cs_10_sphere}
	\caption{Equi-edge grid}
\end{subfigure}
	\caption{(a) Illustration of the resulting gridlines
		for the cubed-sphere equiangular and equi-edge mapping for $N=10$.\label{cs-grids-N10}}
\end{figure}

@@

\section{The conservative transport equation on the cubed-sphere}
\label{adv-2d}
The goal of this Section is to present and solve numerically the conservative transport equation on the cubed-sphere. 
We are going to consider the equi-edge or equiangular cubed-sphere mappings $\Psi_p$ and their respective local coordinates $(x,y)$.
\textcolor{red}{Once again, the metric term is denoted by $\sqrt{\mathfrak{g}}_{\Psi}$ by $\sqrt{\mathfrak{g}}$ for simplicity.}
Following \cite{nair:2010}, the transport model without sources or sinks is considered:
\begin{align}
	\label{transp-rho}
	[\partial_t \rho + \nabla \cdot (\rho \boldsymbol{u})](x,y,p,t)&=0,\\
	\label{transp-phi}
	[\partial_t (\rho \phi) + \nabla \cdot (\rho \phi \boldsymbol{u})](x,y,p,t)&=0,
\end{align}
for $x,y\in[-a,a]$, $t \in [0,T]$, where $T$ is the final time,
\textcolor{red}
{$\boldsymbol{u}$ is the wind, $\mathfrak{u}$ and $\mathfrak{v}$ are the contravariant wind components (Equation \ref{contravariant-wind}),
$\rho$ is the fluid density and $\phi$ is the tracer concentration}.
Additionally, for the transport model, it is assumed that the initial conditions are given as $\rho(x,y,p,0)=\rho_0(x,y,p)$ and $\phi(x,y,p,0)=\phi_0(x,y,p)$, given $\rho_0$ and $\phi_0$. % and that the boundary conditions are related to the cubed-sphere topology.
Equation \eqref{transp-rho} is the continuity equation and Equation \eqref{transp-phi} is the conservative advection equation.
In this framework, $\rho \phi$ represents the tracer density, and the total masses of $\rho$ and $\rho \phi$ are preserved.  Therefore, these quantities are referred to as conserved quantities.

@@

To concluded this Section, the transport model described by Equations \eqref{transp-rho} and \eqref{transp-phi} may be solved using the scheme from Equation \eqref{tr-scheme}, applied considering $q^n=\rho^n$ and $q^n=(\rho \phi)^n$ simultaneously.
In this framework, the tracer concentration is given by $\phi^n = \frac{(\rho \phi)^n}{\rho^n}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical experiments}
\label{num-exp}
\textcolor{red}{
In this Section, the goal is to present simulations of the numerical solution of the transport model, governed by Equations \eqref{transp-rho} and \eqref{transp-phi}, utilizing the FV3 and LT2 schemes. As mentioned earlier, the tracer concentration $\phi$ is advected in the transport model.}
To specify the simulation, including the initial condition for the tracer concentration $\phi_0$, the zonal wind component  denoted by $u_\lambda$, and the meridional wind component denoted by $v_\theta$, need to be defined.
The conversion from these wind components to cubed-sphere contravariant wind components (Equation \eqref{contravariant-wind}) is detailed in \ref{ap-wind}.

@@

\subsection{Rotated zonal wind experiments}
In this Section, the following rotated zonal wind field based on \cite{will:1992} is considered:
\begin{align}
	\label{zonal-wind}
	\begin{cases}
		u_\lambda(\lambda,\theta,t) = u_0(\cos(\theta)\cos(\alpha) + \sin(\theta)\cos(\lambda)\sin(\alpha)),\\
		v_\theta(\lambda,\theta,t) = -u_0\sin(\lambda)\sin(\alpha),
	\end{cases}
\end{align}
where $\lambda \in [-\pi,\pi]$ represents longitude and $\theta \in [-\frac{\pi}{2},\frac{\pi}{2}]$ represents latitude. 
It is easy to see that the wind is divergence free.
The parameter $\alpha$ is the rotation angle. Following \cite{putman:2007}, the rotation angle is set as $\alpha=\frac{\pi}{4}$ so that the wind is oriented with the cube corners. 
Finally, the parameter $u_0$ is defined as $u_0 = \frac{2\pi R}{T}$.
With this choice, the simulation period is 12 days. The solution should converge to the initial condition after this period, enabling us to compute the final error.
Therefore, the temporal evolution of the error may be analyzed.
For an expression of the exact solution in this case and a general initial condition, refer to \cite[Theorem 5.1, p. 155]{brachet:2018}.
The time-step for $N=48$ is given by $\Delta t^{(0)} = 3600$ seconds, leading to a CFL number of approximately 0.95.
